Automating Verification of MPC Security: $\metaprot$ to $\minifed$

### 2023-07-04 popl24/scratch.tex
checkpoint

\begin{lemma}
  \label{lemma-separation}
  The following properties hold:
  \begin{enumerate}
  \item $\sep{\pmf}{Y}{Z}$ iff $\sep{\pmf}{Z}{Y}$
  \item $\vc{\pmf}{x}{y}$ if  $\vc{\pmf}{y}{x}$
  \item $\vc{\pmf}{x}{y}$ and $\vc{\pmf}{y}{z}$ imply $\vc{\pmf}{x}{z}$
  \item $\sep{\pmf}{X}{Y}$ and $\sep{\pmf}{X}{Z}$ imply $\sep{\pmf}{X}{(Y \cup Z)}$
  \item If $\prog_1;\prog_2$ is safe and $\vars(\prog_1) \cap \vars(\prog_2) = \varnothing$
    then $\sep{\progtt(\prog_1;\prog_2)}{\vars(\prog_1)}{\vars(\prog_2)}$.
  \end{enumerate}
\end{lemma}

@@

@@

\begin{lemma}[Substitution]
  \label{lemma-substitution}
  If $\sep{\progtt(\prog_2)}{\{ f \}}{\vdefs(\prog_2)}$ and
  $\prog_1;\prog_2[\be/f]$ is safe with $\vars(\prog_1,\be) \cap
  \vars(\prog_2) = \varnothing$ then
  $\sep{\progtt(\prog_1;\prog_2[\be/f])}{\vdefs(\prog_1)}{\vdefs(\prog_2[\be/f])}$.
\end{lemma}
\begin{proof}
  Suppose on the contrary it was not the case that
  $\sep{\progtt(\prog_1;\prog_2[\be/f])}{\vdefs(\prog_1)}{\vdefs(\prog_2[\be/f])}$.
  Then by Lemma \ref{lemma-sepjoin} there exists $\store_1,\store_2
  \in \runs(\prog_1;\prog_2[\be/f])$ with no $\store \in
  \runs(\prog_1;\prog_2[\be/f])$ such that $\store^1_{\vdefs(\prog_1)}
  \cap \store^2_{\vdefs(\prog_2)} \subseteq \store$.  But also by
  assumption and Lemma \ref{lemma-sepjoin} for any $\beta$ there
  exists $\store' \in \runs(\prog_2)$ with $\{ f \mapsto \beta \} \cap
  \store^2_{\vdefs(\prog_2)} \subseteq \store'$. So in particular, we
  have $\{ f \mapsto \lcod{\store^1_{\vdefs(\prog_1)},\be}{i} \} \cap
  \store^2_{\vdefs(\prog_2)} \subseteq \store'$ for any $i$. This, the
  assumption $\vars(\prog_1,\be) \cap \vars(\prog_2) = \varnothing$,
  and application of Lemma \ref{lemma-presub} leads to the consequence
  that there exists $\store \in \runs(\prog_1;\prog_2[\be/f])$ with
  $\store \supseteq \store^1_{\vdefs(\prog_1)} \cap
  \store^2_{\vdefs(\prog_2)}$ given the assumption $\vars(\prog_1,\be)
  \cap \vars(\prog_2) = \varnothing$, which is a contradiction.
\end{proof}

\begin{lemma}[GMW Gate Composition]
  \label{lemma-composition-gmw}
  Given well-typed $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$ and certified $\mathit{gate}$ where:
  $$
  \begin{array}{lcr}
   v_1 &=& \ttt{\{c1 = v[1,g1out]; c2 = v[2,g1out]\}}\\
   v_2 &=& \ttt{\{c1 = v[1,g2out]; c2 = v[2,g2out]\}}
  \end{array}
  $$
  for some $\ttt{g1}$ and $\ttt{g2}$ and:
  $$
  \config{\prog_1}{E[\mathit{gate}(\ttt{g},v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[\ttt{\{c1 = v[1,gout]; c2 = v[2,gout]\}}]}
  $$
  Then the following postconditions hold:
  \begin{enumerate}
  \item $\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out]\}}}{\ttt{\{v[2,gout]\}}}$
  \item $\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[2,g1out],v[2,g2out]\}}}{\ttt{\{v[1,gout]\}}}$
  \end{enumerate}
\end{lemma}

\begin{proof}
  Let $\prog$ be as defined in Definition \ref{definition-gmwgate-certification}.
  We observe:
  $$
  {\footnotesize
    \begin{array}{c}
      \prog_2 = \\
      \prog[\ttt{v[1,g1out]}/\ttt{flip[1,s11]}][\ttt{v[2,g1out]}/\ttt{flip[2,s12]}][\ttt{v[1,g2out]}/\ttt{flip[1,s21]}][\ttt{v[2,g2out]}/\ttt{flip[2,s22]}]
    \end{array}
  }
  $$
  The assumption of well-typedness of $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$ assures
  that $\ttt{g}$ has not been previously used as a gate identifier, since otherwise the
  views $\ttt{v[1,gout]}$ and $\ttt{v[1,gout]}$ would have been previously defined. This
  ensures $\vars(\prog_1,x) \cap
  \vars(\prog_2) = \varnothing$ for:
  $$x \in \{ \ttt{v[1,g1out]}, \ttt{v[2,g1out]}, \ttt{v[1,g2out]}, \ttt{v[2,g2out]}\}$$
  given condition (iii) in Definition \ref{definition-gmwgate-certification}.
  The result follows by Lemma \ref{lemma-substitution}. 
\end{proof}

\begin{lemma}[GMW Gate Invariance]
  \label{lemma-preservation-gmw}
  Given well-typed $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$ and certified $\mathit{gate}$ where
  for some $\ttt{g1}$ and $\ttt{g2}$, :
  $$
  \begin{array}{lcr}
   v_1 &=& \ttt{\{c1 = v[1,g1out]; c2 = v[2,g1out]\}}\\
   v_2 &=& \ttt{\{c1 = v[1,g2out]; c2 = v[2,g2out]\}}
  \end{array}
  $$
  and where $\iov(\prog_1) = S \cup V$ we have:
  $$
  \sep{\progd(\prog_1)}{S_{\{2\}}}{V_{\{1\}}}
  \quad \text{and} \quad \sep{\progd(\prog_1)}{S_{\{1\}}}{V_{\{2\}}}
  $$
  and:
  $$
  \config{\prog_1}{E[\mathit{gate}(\ttt{g},v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[\ttt{\{c1 = v[1,gout]; c2 = v[2,gout]\}}]}
  $$
  then we have as a postcondition:
  %\begin{enumerate}
  $$\sep{\progd(\prog_1;\prog_2)}{S_{\{2\}}}{(V_{\{1\}} \cup \ttt{\{v[1,gout]\}})}
  \quad \text{and}\quad \sep{\progd(\prog_1;\prog_2)}{S_{\{1\}}}{(V_{\{2\}} \cup \ttt{\{v[2,gout]\}})}$$
  %\end{enumerate}
\end{lemma}

\begin{proof}
By Lemma \ref{lemma-composition-gmw} we have:
$$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out]\}}}{\ttt{\{v[2,gout]\}}}$$
and
$$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[2,g1out],v[2,g2out]\}}}{\ttt{\{v[1,gout]\}}}$$
\end{proof}

>>>>

\begin{lemma}
  \label{lemma-separation}
  The following properties hold:
  \begin{enumerate}
  \item $\sep{\pmf}{Y}{Z}$ iff $\sep{\pmf}{Z}{Y}$
  \item $\vc{\pmf}{x}{y}$ if  $\vc{\pmf}{y}{x}$
  \item $\vc{\pmf}{x}{y}$ and $\vc{\pmf}{y}{z}$ imply $\vc{\pmf}{x}{z}$
  \item ($\sep{\pmf}{X}{Y}$ and $\sep{\pmf}{X}{Z}$) iff $\sep{\pmf}{X}{(Y \cup Z)}$
  %\item If $\prog_1;\prog_2$ is safe and $\vars(\prog_1) \cap \vars(\prog_2) = \varnothing$
  %  then $\sep{\progtt(\prog_1;\prog_2)}{\vars(\prog_1)}{\vars(\prog_2)}$.
  \end{enumerate}
\end{lemma}

%\begin{lemma}[Wire Framing]
%  \label{lemma-wire-framing}
%  If $\sep{\progtt(\prog_2)}{X}{Z}$ and
%  $\vars(\prog_1) \cap \vars(\prog_2) \subseteq X$ then
%  $\sep{\progtt(\prog_1;\prog_2)}{Y}{Z}$ for
%  $Y \subseteq \vars(\prog_1)$.
%\end{lemma}

@@

\begin{lemma}
  \label{lemma-gmw-frame}
  If $\sep{\progtt(\prog_2)}{Y}{Z}$ and $\sep{\progtt(\prog_1)}{X}{Y}$ for
  $X \not\in \vars(\prog_2)$ then $\sep{\progtt(\prog_1;\prog_2)}{X}{Z}$. 
\end{lemma}

@@

\begin{lemma}[Substitution]
  \label{lemma-substitution}
  If $\sep{\progtt(\prog_2)}{\{ f \}}{X}$ and
  $\prog_1;\prog_2[\be/f]$ is safe with $\vars(\prog_1,\be) \cap
  X = \varnothing$ then
  $\sep{\progtt(\prog_1;\prog_2[\be/f])}{\vars(\be)}{X}$.
\end{lemma}
\begin{proof}
  Suppose on the contrary it was not the case that
  $\sep{\progtt(\prog_1;\prog_2[\be/f])}{\vars(\be)}{X}$.
  Then by Lemma \ref{lemma-sepjoin} there exists $\store_1,\store_2
  \in \runs(\prog_1;\prog_2[\be/f])$ with no $\store \in
  \runs(\prog_1;\prog_2[\be/f])$ such that $\store^1_{\vars(\be)}
  \cap \store^2_X \subseteq \store$.  But also by
  assumption and Lemma \ref{lemma-sepjoin} for any $\beta$ there
  exists $\store' \in \runs(\prog_2)$ with $\{ f \mapsto \beta \} \cap
  \store^2_X \subseteq \store'$. So in particular, we
  have $\{ f \mapsto \lcod{\store^1_{\vars(\be)},\be}{i} \} \cap
  \store^2_X \subseteq \store'$ for any $i$. This, the
  assumption $\vars(\prog_1,\be) \cap \vars(\prog_2) = \varnothing$,
  and application of Lemma \ref{lemma-presub} leads to the consequence
  that there exists $\store \in \runs(\prog_1;\prog_2[\be/f])$ with
  $\store \supseteq \store^1_{\vars(\be)} \cap
  \store^2_X$ given the assumption $\vars(\prog_1,\be)
  \cap \vars(\prog_2) = \varnothing$, which is a contradiction.
\end{proof}

\begin{lemma}[GMW Gate Invariant]
  \label{lemma-preservation-gmw}
  Given well-typed $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$ and certified $\mathit{gate}$ where
  for some $\ttt{g1}$ and $\ttt{g2}$, :
  $$
  \begin{array}{lcr}
   v_1 &=& \ttt{\{c1 = v[1,g1out]; c2 = v[2,g1out]\}}\\
   v_2 &=& \ttt{\{c1 = v[1,g2out]; c2 = v[2,g2out]\}}
  \end{array}
  $$
  and let:
  $$
  \config{\prog_1}{E[\mathit{gate}(\ttt{g},v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[\ttt{\{c1 = v[1,gout]; c2 = v[2,gout]\}}]}
  $$
  If we have the following precondition, where $\iov(\prog_1) = S \cup V$:
  $$
  \sep{\progd(\prog_1)}{(S_{\{2\}} \cup V_{\{2\}})}{(S_{\{1\}} \cup V_{\{1\}})}
  $$
  then we have as a postcondition:
  $$
  \sep{\progd(\prog_1)}{(S_{\{2\}} \cup V_{\{2\}} \cup \{ \ttt{v[2,gout]} \})}{(S_{\{1\}} \cup V_{\{1\}} \cup \{ \ttt{v[1,gout]} \})}
  $$
\end{lemma}

\begin{proof}
  Let $\prog$ be as defined in Definition \ref{definition-gmwgate-certification}.
  We observe:
  $$
  {\footnotesize
    \begin{array}{c}
      \prog_2 = \\
      \prog[\ttt{v[1,g1out]}/\ttt{flip[1,s11]}][\ttt{v[2,g1out]}/\ttt{flip[2,s12]}][\ttt{v[1,g2out]}/\ttt{flip[1,s21]}][\ttt{v[2,g2out]}/\ttt{flip[2,s22]}]
    \end{array}
  }
  $$ Now, the assumption of well-typedness of
  $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$ assures that $\ttt{g}$
  has not been previously used as a gate identifier in $\prog_1$,
  since otherwise the views $\ttt{v[1,gout]}$ and $\ttt{v[2,gout]}$
  would have been previously defined. This ensures $\vars(\prog_1,x)
  \cap \vars(\prog_2) = \varnothing$ for:
  $$x \in \{ \ttt{v[1,g1out]}, \ttt{v[2,g1out]}, \ttt{v[1,g2out]}, \ttt{v[2,g2out]}\}$$
  given condition (iii) in Definition \ref{definition-gmwgate-certification}.
  Thus by Lemma \ref{lemma-substitution} we have:
  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out]\}}}{\ttt{\{v[2,gout]\}}}$$
  and
  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[2,g1out],v[2,g2out]\}}}{\ttt{\{v[1,gout]\}}}$$
  Since $\vars(\prog_1) \cap \vars(\prog_2)$ are only the wire inputs to $\ttt{g}$,
  therefore by assumption and Lemma \ref{lemma-gmw-frame}:
  $$\sep{\progtt(\prog_1;\prog_2)}{(S_{\{1\}} \cup V_{\{1\}})}{\ttt{\{v[2,gout]\}}}$$
  and also:
  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,g1out],v[1,g2out]\}}}{(S_{\{2\}}\cup V_{\{2\}})}$$
  etc., so the result follows. 
\end{proof}
